Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-05-11 Documents/M1/dgM1Curvilinear.tex
Completed Ab/Em part in chapter 6. Isoenergetic scattering needs to be added.

\ee{Ran write about tabulated B85 absorption and scattering opacities.}

>>>>

%\ee{Ran write about tabulated B85 absorption and scattering opacities.}

Given by \citet{bruenn_1985}, the absorption opacity of neutrino $\chi$ 
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \chi(\epsilonNu,\bU) + \eta(\epsilonNu,\bU), 
\end{equation}
where $\eta(\epsilonNu,\bU) $ is the neutrino emissivity
\begin{equation}
\eta = \eta_{nucleon} + \eta_{nuclear},
\end{equation}  
and $ \chi(\epsilonNu,\bU)$ is the neutrino absorptivity, which is the inverse of neutrino absorption mean free path.
They have the following relation:
\begin{equation}
\chi(\epsilonNu,\bU) = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} \eta(\epsilonNu,\bU). 
\end{equation}

Therefore,
\begin{equation}
\chi = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} (\eta_{nucleon} + \eta_{nuclear} ) = \chi_{nucleon} + \chi_{nuclear}.
\end{equation}
The two terms of electron-neutrino absorption on nucleons are
 \begin{align}
    \chi_{nucleon}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} \eta_{np}(g_V^2+3g_A^2)[1-F_e(\varepsilon+Q)](\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonabsorption} ,\\
    \eta_{nucleon}(\epsilonNu,\bU) &= \dfrac{G^2}{\pi}\eta_{pn}(g_V^2+3g_A^2)F_e(\varepsilon+Q)(\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonemissivity} ,
\end{align}
and the two terms of electron-neutrino absorption on nuclear are
\begin{align}
    \chi_{nuclear}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} n_A e^{(\mu_n -\mu_p -Q')/k_B T} g^2_A \dfrac{2}{7} N_p(Z) N_h(N)[1-F_e(\varepsilon+Q')](\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearabsorption} ,\\
    \eta_{nuclear}(\epsilonNu,\bU) &=\dfrac{G^2}{\pi} n_A g^2_A \dfrac{2}{7} N_p(Z) N_h(N)F_e(\varepsilon+Q')(\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearemissivity},
\end{align} 
where $g_V = 1$ and $g_A =1.23$ are form factors resulting from virtual strong-interaction processes, $M_e = 0.511$ MeV is the electron rest energy, $Q = 1.2935$ MeV is the difference between the neutron and proton rest energies, $\mu_n$ and $\mu_p$ are the neutron and proton chemical potential, respectively. And $Q'$ is defined by $Q' = \mu_n -\mu_p + \Delta$, where $\Delta=3$MeV is the energy of the $1f_{5/2}$ level above the ground state.

In Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity}, $F_e(E)$ is the Fermi-Dirac distribution function:
\begin{align}
F_e(E) = \dfrac{1}{e^{(E-\mu_e)/k_B T} + 1},
\end{align}
where $\mu_e$ is the electron chemical potential.

In Eq.~\eqref{eq:nucleonabsorption} and ~\eqref{eq:nucleonemissivity}, $\eta_{np}$ and $\eta_{pn}$ take into account the nucleon final-state blocking and equal to $n_n$ and $n_p$, the nucleon number density, when the nucleons are nondegenerate, respectively.
\begin{align}
n_p & =  \dfrac{X_p \rho}{m_p} \\
n_n & =  \dfrac{X_n \rho}{m_n},
\end{align}
where $X_p$ and $X_n$ are the proton and neutron fractions, respectively.

In Eq.~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity}, $n_A$ is the nuclei number density and $n_A = \dfrac{\rho X_H}{m_B A}$. The function $N_p$ is the number of protons in the single-particle $1f_{7/2}$ level and $N_h$ is the number of neutron holes in the single-particle $1f_{5/2}$ level. As a crude estimate of $N_p$ and $N_h$, a zero-order shell model predicts that
\[   
N_p(Z) = 
     \begin{cases}
       0, &\quad Z<20\\
       Z-20, &\quad 20<Z<28 \\
       8, &\quad Z>28,\\
     \end{cases}
\]
and
\[   
N_h(N) = 
     \begin{cases}
       6, &\quad N<34\\
       40-N, &\quad 34<N<40 \\
       0, &\quad N>40.\\
     \end{cases}
\]

The factor $G^2$ in Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} is the Fermi constant [$G=(1.015\pm 0.03)\time 10^5 M_p^{-2}$] with the value
\begin{equation}
	G^2 = \left[ \dfrac{1}{(\hbar c)^4} \right] G^2 = 5.18 \times 10^{-44} MeV^{-2}cm^2, 
\end{equation}
giving $\chi$ and $\eta$ the required units of $cm^{-1}$.

The sum of Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} gives the explicit expression of the absorption opacity of neutrino $\chi$ .

The total opacity $\kappa(\epsilonNu,\bU)=\chi(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, and $\sigma(\epsilonNu,\bU)$ is the scattering opacities. The scattering opacities $\sigma(\epsilonNu,\bU)$ contains isoenergetic scattering opacities and have a integral form in the Boltzmann equation: if we include emission, absorption, and isoenergetic scattering of neutrino by nucleons and nuclei, the simplified Boltzmann equation is
\begin{align}
\dfrac{\partial f(\epsilonNu,\bU)}{\partial t} = \eta(\epsilonNu,\bU) - \tilde{\chi}(\epsilonNu,\bU)f(\epsilonNu,\bU) & + \dfrac{1}{h^3} \epsilonNu^2 \int d\mu^\prime_0 R_{IS}( \mu^\prime_0, \mu_0,\epsilonNu, \bU) f(\mu^\prime_0, \epsilonNu,\bU) \nonumber \\
	& - \dfrac{1}{h^3} \epsilonNu^2 f(\epsilonNu,\bU)\int d\mu^\prime_0 R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) .
\end{align}

The last two terms are contributions from isoenergetic scattering of neutrino by nucleons and nuclei. The isoenergetic scattering kernel $R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)$ has the following relation:
\begin{align}
R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) & = R_{IS}^{nucleons}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) +  R_{IS}^{nuclei}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)\\
& = \int_{0}^{2\pi}d
\end{align}
\clearpage
